---
title: "Untitled"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, eval = T, results = 'hide',
                      fig.align = 'center', fig.width = 8, fig.height = 8)
knitr::opts_knit$set(root.dir = here::here())
```
```{r}
plot_theme <- theme(plot.title = element_text(family = "Helvetica", face = "bold", hjust = 0.5, size = 14),
                    axis.title = element_text(family = "Helvetica", face = "bold", hjust = 0.5, size = 12),
                    axis.text.y = element_text(family = "Helvetica", hjust = 0.5, size = 11),
                    axis.text.x = element_text(family = "Helvetica", hjust = 0.5, size = 11, angle = 90),
                    legend.position = "right",
                    legend.title = element_text(family = "Helvetica", face = "bold", vjust = 0.85, size = 12),
                    legend.text = element_text(family = "Helvetica", hjust = 0.5, vjust = 0.5, size = 10),
                    axis.line = element_line(color = "#000000", size = 1))
```

# Overview 
From the discete-loci discrete-time scructured Wright-Fisher model for _P. falciparum_ genetics, I ran 10 iterations for "base" simulations (`migration = 0` and/or `COI = 1` in all mosquito guts [resulting in limited recombination]) and 100 iterations for "interesting" simulations. Simulations are essentially identical to those used in Verity, Aydemir, Brazeau _et al._ 2020, where the migration paramter M can be: _0, 0.25, 0.5, or 1_ and deme size is  from (N) on a logarithmic-base 10 scale: _10, 16, 25, 40, 63, 100, 158, 251, 398, 631, 1000_.



```{r}
#......................
# imports
#......................
library(tidyverse)
library(HMMERTIME)
```
```{r}
#......................
# read in
#......................
hmmret <- readRDS("results/tidied_results/hmmertime_results_from_dtdlswfsims.RDS")
#......................
# factorize for plotting
#......................
coilvls <- sort(unique(hmmret$mean_coi))
hmmret <- hmmret %>% 
  dplyr::mutate(mean_coi_fact = factor(mean_coi, 
                                       levels = coilvls,
                                       labels = c("Mean COI: 1", "Mean COI: 2", "Mean COI: 3")),
                m_fact = factor(m, 
                                levels = c("0", "0.25", "0.5", "1"),
                                labels = c("Migr: 0", "Migr: 0.25", "Migr: 0.5", "Migr: 1")),
                logN = log10(N)) 

#......................
# put in tidy format for easier plotting
#......................
hmmret <- hmmret %>% 
  dplyr::mutate(true_coi = purrr::map(trueIBD, "eff_coi"),
                true_coi1 = purrr::map_dbl(true_coi, function(x) as.numeric( stringr::str_split(x, " ", n = 2)[[1]] )),
                true_coi2 = purrr::map_dbl(true_coi, function(x) as.numeric( stringr::str_split(x, " ", n = 2)[[2]] )),
                pairwiseIBD = purrr::map_dbl(trueIBD, "pairwiseIBD", ),
                effpairwiseIBD = purrr::map_dbl(trueIBD, "effpairwiseIBD"),
                withinIBD_smpl1 = purrr::map_dbl(trueIBD, "withinIBD_smpl1"),
                withinIBD_smpl2 = purrr::map_dbl(trueIBD, "withinIBD_smpl2"),
                hmmertimequantdf = purrr::map(hmmertimequants, function(x){
                  ret <- as.data.frame(x)
                  colnames(ret) <- paste0("Q", colnames(ret))
                  colnames(ret) <- sub("%", "", colnames(ret))
                  ret$param <- rownames(ret)
                  return(ret)
                })) %>% 
  tidyr::unnest(cols = "hmmertimequantdf")
```

## Compare COIs
```{r, results = 'asis'}
COI1plotObj <- hmmret %>% 
  dplyr::filter(param == "m1") %>% 
  ggplot() + 
  geom_pointrange(aes(x = true_coi1, y = Q50, ymin = Q2.5, ymax = Q97.5, 
                      color = logN), alpha = 0.5) +
  scale_color_viridis_c("Log Ne") +
  geom_abline(slope = 1, intercept = 0, color = "#3C3B6E") + 
  geom_smooth(aes(x = true_coi1, y = Q50), method = "lm", color = "#B22234") + 
  facet_grid(mean_coi_fact ~ m_fact) + 
  ylab("Inferred COI") + xlab("True COI") +
  ggtitle("COI of First Sample") + 
  labs(caption = "Expected is in Blue; Red is slope of inferred") +
  plot_theme 


COI2plotObj <- hmmret %>% 
  dplyr::filter(param == "m2") %>% 
  ggplot() + 
  geom_pointrange(aes(x = true_coi2, y = Q50, ymin = Q2.5, ymax = Q97.5, 
                      color = logN), alpha = 0.5) +
  scale_color_viridis_c("Log Ne") +
  geom_abline(slope = 1, intercept = 0, color = "#3C3B6E") + 
  geom_smooth(aes(x = true_coi2, y = Q50), method = "lm", color = "#B22234") + 
  facet_grid(mean_coi_fact ~ m_fact) + 
  ylab("Inferred COI") + xlab("True COI") +
  ggtitle("COI of Second Sample") + 
  labs(caption = "Expected is in Blue; Red is slope of inferred") +
  plot_theme 

cowplot::plot_grid(COI1plotObj, COI2plotObj, nrow = 1, align = "h")


```


## Compare IBD Measures

```{r, results = 'asis'}
btwnIBDplotObj <- hmmret %>% 
  dplyr::filter(param == "f_ind") %>% 
  ggplot() + 
  geom_pointrange(aes(x = pairwiseIBD, y = Q50, ymin = Q2.5, ymax = Q97.5, 
                      color = logN), alpha = 0.5) +
  scale_color_viridis_c("Log Ne") +
  geom_abline(slope = 1, intercept = 0, color = "#3C3B6E") + 
  geom_smooth(aes(x = pairwiseIBD, y = Q50), method = "lm", color = "#B22234") + 
  facet_grid(mean_coi_fact ~ m_fact) + 
  ylab("Inferred Individal IBD") + xlab("True Individaul IBD") +
  ggtitle("Between Sample IBD") + 
  labs(caption = "Expected is in Blue; Red is slope of inferred") +
  plot_theme 
```
